---
title: "SUPPLEMENTARY DATA"
subtitle: "Co-occurrence and character convergence in two Neotropical bats"
author:
  - Luis Darcy Verde Arregoitia
  - Giovani Hernández-Canchola
  - Luca Santini
  - Manuel Schweizer
  - Martín Y. Cabrera-Garrido
  - Livia S. León-Paniagua
output:
  pdf_document: default
  html_document:
    df_print: paged
  word_document: default
---
*Journal of Mammalogy*  
  


```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  comment = "#>"
)
library(tidyverse)
library(here)
library(dplyr)
library(purrr)
library(here)
library(forcats)
library(knitr)

```

```{r LMM_filtered, echo = FALSE, include=FALSE }

# read the csv files
# read the exported effects for reporting
AlleffectsFilt <- 
  list.files(here("lmmEffects","filtered"), pattern = "output.csv$", full.names = TRUE) %>% 
  map_df(read_csv) %>% select(-est.)

# relevel and arrange
AlleffectsFilt$body.measurement <-  fct_relevel(AlleffectsFilt$body.measurement,"forearm length","body mass","skull isosize","head length")
AlleffectsFilt <- AlleffectsFilt %>% arrange(body.measurement)

# rename column
AlleffectsFilt <- AlleffectsFilt %>% rename(variable="term",SE=std.error)
# round ICC
AlleffectsFilt$ICC <- round(AlleffectsFilt$ICC,3)

# for reporting
# drop SE and ICC, rename CI columns
AlleffectsFilt <- AlleffectsFilt %>% select(-SE,-ICC) %>% rename("CIlower"=lower,"CIupper"="upper")

# number formatting
AlleffectsFilt <-  AlleffectsFilt %>% mutate_if(is.numeric, funs(formatC(.,format="f",digits = 2))) 

# merge estimate and CIs to save space

AlleffectsFilt <-
  AlleffectsFilt %>% mutate(Estimate=paste0(estimate," ","(",CIlower," ","-"," ",CIupper,")")) %>% select(Model=body.measurement,Variable=variable,'Estimate (95CI)'=Estimate) 
AlleffectsFilt <-
  AlleffectsFilt %>% 
  mutate(Variable = recode(Variable,
                           "specieshondurensis"= "species (hondurensis)",
                           "speciesparvidens"  = "species (parvidens)",
                           "sexM" = "sex (male)",
                           "specieshondurensis:sympatryboth" = "species (hondurensis): occurrence (sympatry)",
                           "speciesparvidens:sympatryboth" = "species (parvidens): occurrence (sympatry)"))

# for presentation purposes
AlleffectsFilt <-
AlleffectsFilt %>% mutate(Model=as.character(Model))
AlleffectsFilt <-
AlleffectsFilt %>% mutate(Model=ifelse(duplicated(.[["Model"]]),"",Model))
```

**Supplementary Data S1.** Parameter estimates with upper and lower 95% confidence intervals for the linear mixed models fit separately for each body measurement, excluding localities with only one specimen. 

```{r, echo=FALSE}
SLMM <- kable(AlleffectsFilt,align = c('l','l','r'))
SLMM
```


```{r ANOVA_filtered, echo = FALSE, include = FALSE}
# make all the csv outputs

list.files(here("filtered"), pattern = "LMM", full.names = TRUE) %>%
  map(source)

# reporting LMM anova, filtered

# read the exported ANOVA tables for reporting
AllaovFilt <-
  list.files(here("lmmEffects", "filtered"),
             pattern = "^aov",
             full.names = TRUE) %>%
  map_df(read.csv)

# relevel and arrange
AllaovFilt$measurement <-
  fct_relevel(AllaovFilt$measurement,
              "forearm length",
              "body mass",
              "skull isosize",
              "head length")
AllaovFilt <- AllaovFilt %>% arrange(measurement) %>% select(-df)

# format numbers
AllaovFilt <-
  AllaovFilt %>% mutate_if(is.numeric, funs(formatC(., format = "f", digits = 3)))

# rename column
AllaovFilt <- AllaovFilt %>% rename(model = measurement)

AllaovFilt <-
  AllaovFilt %>% mutate(term = dplyr::recode(term, "species:sympatry" =
                                               "species:occurrence"))
# formatting
AllaovFilt <-
AllaovFilt %>% mutate(Model=as.character(model)) %>% select(-model)
AllaovFilt <-
AllaovFilt %>% mutate(Model=ifelse(duplicated(.[["Model"]]),"",Model)) %>% select(Model,everything())%>% mutate(p.value = case_when(p.value == "0.000" ~ "<0.001", TRUE ~p.value))


```

\newpage

**Supplementary Data S2.** ANOVA results (Wald Chi square test, one degree of freedom) for the linear mixed models fit separately for each body measurement, excluding localities with only one specimen. 
```{r, echo=FALSE}
Saov <- kable(AllaovFilt, align = c('l','l','r','r'))
Saov
```

```{r FigS1, echo = FALSE, include = FALSE}
# sympatry effects plots with no singleton localities
library(dplyr)
library(purrr)
library(here)
library(cowplot)
library(ggplot2)
library(extrafont)
library(hrbrthemes)

# read the exported effects for plotting
AlleffectsFilt <- 
list.files(here("lmmEffects","filtered"), pattern = ".csv$", full.names = TRUE) %>% 
  map_df(read.csv)

# cowplot settings
theme_set(theme_cowplot(font_family = "Roboto",font_size = 12))


# object for each measurement  
# forearm
fleffFilt <- 
AlleffectsFilt %>% filter(measurement=="forearm") %>%
ggplot(aes(x=species, y=fit, pch=species,fill=species),color="black") + 
  geom_pointrange(aes(ymin=lower, ymax=upper),size=0.9)+
  facet_wrap(~sympatry,labeller=labeller(sympatry = c(allop="allopatry",both="sympatry")))+
  ylab("forearm length (mm)")+xlab("")+
  scale_shape_manual(values = c(21, 23))+
  scale_fill_manual(values = c("#335EAD", "#EEBB33"))+
  theme(legend.position="none",axis.text.x = element_blank(),
        strip.background = element_rect(fill ="#FDFCFF"))

# body mass
bmeffFilt <- 
AlleffectsFilt %>% filter(measurement=="bodyMass") %>%
  ggplot(aes(x=species, y=fit, pch=species,fill=species),color="black") + 
  geom_pointrange(aes(ymin=lower, ymax=upper),size=0.9)+
  facet_wrap(~sympatry,labeller=labeller(sympatry = c(allop="allopatry",both="sympatry")))+
  ylab("body mass (g)")+xlab("")+
  scale_shape_manual(values = c(21, 23))+
  scale_fill_manual(values = c("#335EAD", "#EEBB33"))+
  theme(legend.position="none",axis.text.x = element_blank(),
        strip.background = element_rect(fill ="#FDFCFF"))

# skull isosize
ISeffFilt <- 
AlleffectsFilt %>% filter(measurement=="isosize") %>%
  ggplot(aes(x=species, y=fit, pch=species,fill=species),color="black") + 
  geom_pointrange(aes(ymin=lower, ymax=upper),size=0.9)+
  facet_wrap(~sympatry,labeller=labeller(sympatry = c(allop="allopatry",both="sympatry")))+
  ylab("skull isosize")+xlab("")+
  scale_shape_manual(values = c(21, 23))+
  scale_fill_manual(values = c("#335EAD", "#EEBB33"))+
  theme(legend.position="none",axis.text.x = element_blank(),
        strip.background = element_rect(fill ="#FDFCFF"))

# head length
CBLeffFilt <- 
AlleffectsFilt %>% filter(measurement=="CBL") %>%
  ggplot(aes(x=species, y=fit, pch=species,fill=species),color="black") + 
  geom_pointrange(aes(ymin=lower, ymax=upper),size=0.9)+
  facet_wrap(~sympatry,labeller=labeller(sympatry = c(allop="allopatry",both="sympatry")))+
  ylab("head length (mm)")+xlab("")+
  scale_shape_manual(values = c(21, 23))+
  scale_fill_manual(values = c("#335EAD", "#EEBB33"))+
  theme(legend.position="none",axis.text.x = element_blank(),
        strip.background = element_rect(fill ="#FDFCFF"))


######## side by side plots 
# dummy plot for shared legend
# hack to separate the elements
legplot <- 
  AlleffectsFilt %>% filter(measurement=="CBL") %>%
  ggplot(aes(x=species, y=fit, pch=species,fill=species),color="black") + 
  geom_pointrange(aes(ymin=lower, ymax=upper),size=0.9)+
  facet_wrap(~sympatry,labeller=labeller(sympatry = c(allop="allopatry",both="sympatry")))+
  scale_shape_manual(values = c(21, 23),labels=c("S. hondurensis   ","S. parvidens"))+
  scale_fill_manual(values = c("#335EAD", "#EEBB33"),labels=c("S. hondurensis   ","S. parvidens"))+
  theme(legend.text = element_text(face = "italic"),legend.title=element_blank())

## extract the legend 
sharedLegend <- get_legend(legplot+theme(legend.position="bottom",legend.justification="center"))

# fig 3
figS3 <- plot_grid(fleffFilt,bmeffFilt,ISeffFilt,CBLeffFilt,nrow = 1)
figS3

## Figure 3
# add the legend to the row we made earlier. Give it one-third of the width of one plot (via rel_widths).
figS3L <- plot_grid(figS3, sharedLegend,ncol = 1, rel_heights = c(1.1, .08))
# draw
figS3L
# shared species 'label'
figS3Ls <- add_sub(figS3L, "species",vjust = -4.2,size=12)
# draw
figS3Lsd <- ggdraw(figS3Ls)

```


**Supplementary Data S3.** Estimated values of the four body measurements for each species when in sympatry or allopatry (excluding localities with only one specimen), holding the effects of sex and environmental variables constant. Vertical lines show upper and lower 95% confidence intervals.



```{r drawFig, fig.width=8.1, fig.height=3.5, echo=FALSE}
figS3Lsd
```


\newpage

```{r LMM_clustered, echo = FALSE, include=FALSE }
# read the csv files
# read the exported effects for reporting
AlleffectsClust <- 
  list.files(here("lmmEffects","clustered"), pattern = "output.csv$", full.names = TRUE) %>% 
  map_df(read_csv) %>% select(-est.)

# relevel and arrange
AlleffectsClust$body.measurement <-  fct_relevel(AlleffectsClust$body.measurement,"forearm length","body mass","skull isosize","head length")
AlleffectsClust <- AlleffectsClust %>% arrange(body.measurement)

# rename column
AlleffectsClust <- AlleffectsClust %>% rename(variable="term",SE=std.error)

# round ICC
AlleffectsClust$ICC <- round(AlleffectsClust$ICC,3)

# for reporting
# drop SE and ICC, rename CI columns
AlleffectsClust <- AlleffectsClust %>% select(-SE,-ICC) %>% rename("CIlower"=lower,"CIupper"="upper")

# number formatting
AlleffectsClust <-  AlleffectsClust %>% mutate_if(is.numeric, funs(formatC(.,format="f",digits = 2))) 

# merge estimate and CIs to save space

AlleffectsClust <-
  AlleffectsClust %>% mutate(Estimate=paste0(estimate," ","(",CIlower," ","-"," ",CIupper,")")) %>% dplyr::select(Model=body.measurement,Variable=variable,'Estimate (95CI)'=Estimate) 
AlleffectsClust <-
  AlleffectsClust %>% 
  mutate(Variable = dplyr::recode(Variable,
                           "specieshondurensis"= "species (hondurensis)",
                           "speciesparvidens"  = "species (parvidens)",
                           "sexM" = "sex (male)",
                           "specieshondurensis:sympatryboth" = "species (hondurensis): occurrence (sympatry)",
                           "speciesparvidens:sympatryboth" = "species (parvidens): occurrence (sympatry)"))

# for presentation purposes
AlleffectsClust <-
  AlleffectsClust %>% mutate(Model=as.character(Model))
AlleffectsClust <-
  AlleffectsClust %>% mutate(Model=ifelse(duplicated(.[["Model"]]),"",Model))
```

**Supplementary Data S4.** Parameter estimates with upper and lower 95% confidence intervals for the linear mixed models fit separately for each body measurement, defining allopatry and sympatry using spatial clustering. 

```{r, echo=FALSE}
SLMMCl <- kable(AlleffectsClust,align = c('l','l','r'))
SLMMCl
```


```{r ANOVA_clustered, echo = FALSE, include = FALSE}
# make all the csv outputs

#list.files(here("clustered"), pattern = "LMM", full.names = TRUE) %>%
# map(source)

# reporting LMM anova, clustered

# read the exported ANOVA tables for reporting
AllaovClust <-
  list.files(here("lmmEffects", "clustered"),
             pattern = "^aov",
             full.names = TRUE) %>%
  map_df(read.csv)

# relevel and arrange
AllaovClust$measurement <-
  fct_relevel(AllaovClust$measurement,
              "forearm length",
              "body mass",
              "skull isosize",
              "head length")
AllaovClust <- AllaovClust %>% arrange(measurement) %>% select(-df)

# format numbers
AllaovClust <-
  AllaovClust %>% mutate_if(is.numeric, funs(formatC(., format = "f", digits = 3)))

# rename column
AllaovClust <- AllaovClust %>% rename(model = measurement)

AllaovClust <-
  AllaovClust %>% mutate(term = dplyr::recode(term, "species:sympatry" =
                                                "species:occurrence"))
# formatting
AllaovClust <-
  AllaovClust %>% mutate(Model=as.character(model)) %>% select(-model)
AllaovClust <-
  AllaovClust %>% mutate(Model=ifelse(duplicated(.[["Model"]]),"",Model)) %>% select(Model,everything())%>% mutate(p.value = case_when(p.value == "0.000" ~ "<0.001", TRUE ~p.value))


```

\newpage

**Supplementary Data S5.** ANOVA results (Wald Chi square test, one degree of freedom) for the linear mixed models fit separately for each body measurement, defining allopatry and sympatry using spatial clustering. 
```{r, echo=FALSE}
SaovCl <- kable(AllaovClust,align = c('l','l','r','r'))
SaovCl
```

```{r FigS2, echo = FALSE, include = FALSE}
# sympatry effects plots with clustering approach
library(purrr)
library(here)
library(ggplot2)
library(cowplot)
library(extrafont)


# read the exported effects for plotting
AlleffectsClust <- 
  list.files(here("lmmEffects","clustered"), pattern = ".csv$", full.names = TRUE) %>% 
  map_df(read.csv)

# cowplot settings
theme_set(theme_cowplot(font_family = "Roboto",font_size = 12))

# object for each measurement  
# forearm
fleffClust <- 
  AlleffectsClust %>% filter(measurement=="forearm") %>%
  ggplot(aes(x=species, y=fit, pch=species,fill=species),color="black") + 
  geom_pointrange(aes(ymin=lower, ymax=upper),size=0.9)+
  facet_wrap(~sympatry,labeller=labeller(sympatry = c(allop="allopatry",both="sympatry")))+
  ylab("forearm length (mm)")+xlab("")+
  scale_shape_manual(values = c(21, 23))+
  scale_fill_manual(values = c("#335EAD", "#EEBB33"))+
  theme(legend.position="none",axis.text.x = element_blank(),
        strip.background = element_rect(fill ="#FDFCFF"))

# body mass
bmeffClust <- 
  AlleffectsClust %>% filter(measurement=="bodyMass") %>%
  ggplot(aes(x=species, y=fit, pch=species,fill=species),color="black") + 
  geom_pointrange(aes(ymin=lower, ymax=upper),size=0.9)+
  facet_wrap(~sympatry,labeller=labeller(sympatry = c(allop="allopatry",both="sympatry")))+
  ylab("body mass (g)")+xlab("")+
  scale_shape_manual(values = c(21, 23))+
  scale_fill_manual(values = c("#335EAD", "#EEBB33"))+
  theme(legend.position="none",axis.text.x = element_blank(),
        strip.background = element_rect(fill ="#FDFCFF"))

# skull isosize
ISeffClust <- 
  AlleffectsClust %>% filter(measurement=="isosize") %>%
  ggplot(aes(x=species, y=fit, pch=species,fill=species),color="black") + 
  geom_pointrange(aes(ymin=lower, ymax=upper),size=0.9)+
  facet_wrap(~sympatry,labeller=labeller(sympatry = c(allop="allopatry",both="sympatry")))+
  ylab("skull isosize")+xlab("")+
  scale_shape_manual(values = c(21, 23))+
  scale_fill_manual(values = c("#335EAD", "#EEBB33"))+
  theme(legend.position="none",axis.text.x = element_blank(),
        strip.background = element_rect(fill ="#FDFCFF"))

# head length
CBLeffClust <- 
  AlleffectsClust %>% filter(measurement=="CBL") %>%
  ggplot(aes(x=species, y=fit, pch=species,fill=species),color="black") + 
  geom_pointrange(aes(ymin=lower, ymax=upper),size=0.9)+
  facet_wrap(~sympatry,labeller=labeller(sympatry = c(allop="allopatry",both="sympatry")))+
  ylab("head length (mm)")+xlab("")+
  scale_shape_manual(values = c(21, 23))+
  scale_fill_manual(values = c("#335EAD", "#EEBB33"))+
  theme(legend.position="none",axis.text.x = element_blank(),
        strip.background = element_rect(fill ="#FDFCFF"))


######## side by side plots 
# dummy plot for shared legend
# hack to separate the elements
legplot <- 
  AlleffectsClust %>% filter(measurement=="CBL") %>%
  ggplot(aes(x=species, y=fit, pch=species,fill=species),color="black") + 
  geom_pointrange(aes(ymin=lower, ymax=upper),size=0.9)+
  facet_wrap(~sympatry,labeller=labeller(sympatry = c(allop="allopatry",both="sympatry")))+
  scale_shape_manual(values = c(21, 23),labels=c("S. hondurensis   ","S. parvidens"))+
  scale_fill_manual(values = c("#335EAD", "#EEBB33"),labels=c("S. hondurensis   ","S. parvidens"))+
  theme(legend.text = element_text(face = "italic"),legend.title=element_blank())

## extract the legend 
sharedLegend <- get_legend(legplot+theme(legend.position="bottom",legend.justification="center"))

# fig S2
figS2 <- plot_grid(fleffClust,bmeffClust,ISeffClust,CBLeffClust,nrow = 1)
figS2

## Figure S2
# add the legend to the row we made earlier. Give it one-third of the width of one plot (via rel_widths).
figS2L <- plot_grid(figS2, sharedLegend,ncol = 1, rel_heights = c(1.1, .08))
# draw
figS2L
# shared species 'label'
figS2Ls <- add_sub(figS2L, "species",vjust = -4.2,size=12)
# draw
figS2Lsd <- ggdraw(figS2Ls)

```


**Supplementary Data S6.** Estimated values of the four body measurements for each species when in sympatry or allopatry (defining occurrence patterns using spatial clusering), holding the effects of sex and environmental variables constant. Vertical lines show upper and lower 95% confidence intervals.



```{r drawFigS2, fig.width=8.1, fig.height=3.5, echo=FALSE}
figS2Lsd
```

  
> Density-based clusters were obtained using the R library 'dbscan', an epsilon value of 4ooo m and a minimum of four points for each cluser.

\newpage

**Supplementary Data S7.** Test statistics and p-values for modified signed-likelihood ratio tests (M.SLRT) for equality of coefficients of variation by species in relation to their co-occurrence.  

```{r cv tests, echo=FALSE}
cvtestsResraw <- read.csv(here("variability_tests","vartests.csv"),stringsAsFactors = FALSE)
# format numbers
cvtestsRes <- cvtestsResraw %>% mutate_if(is.numeric, funs(formatC(., format = "f", digits = 3)))
cvoutkable <- kable(cvtestsRes,align = c('l','l','r','r'))
cvoutkable
```

```{r specimen list setup, echo=FALSE, include=FALSE}
# external measurements and environment
sturnfull <- read.csv(here("data","flwbiocl.csv"),stringsAsFactors = FALSE, encoding = "UTF-8") %>%  mutate(spIDs=if_else(!is.na(ncat),as.character(ncat),collector))
# read the skull data
batskulls <- read.csv(here("data","cranBiocl.csv"),stringsAsFactors = FALSE, fileEncoding = "UTF-8")
pander::panderOptions("p.wrap",'')
```

**Supplementary Data S8.** Specimens examined. All specimens are housed in the Mammalogy Collection, Museo de Zoología "Alfonso L. Herrera", Facultad de Ciencias, Universidad Nacional Autónoma de México. Listings show catalog numbers or unique collector IDs. Morphology and location data for each specimen are available in the online repository (https://github.com/luisDVA/SturniraMacroecology)

Skulls examined  
  
```{r echo=FALSE}
pander::pander(batskulls$ncat)
```

Skins examined  
  
```{r echo=FALSE}
pander::pander(sturnfull$spIDs)
```